---
title: "Counting graves"
format: html
---

First things first. We need to load the libraries we'll need.

```{r}
library(tidyverse)
library(sf)
library(stars)
library(lubridate)
library(terra)
library(tidyterra)
```

## 26 May 2022

Load our image.

```{r}
site_2_20220526 <- rast("/Users/a0161776/Documents/nicar/10400100757E2A00 3.tif")
```

We can display the image we just loaded using ggplot and tidyterra.

```{r}
site_2_20220526 |>
    ggplot() +
    geom_spatraster_rgb(data=site_2_20220526, r=1, g=2, b=3)
```

So let's crop this big image.

```{r}
crop_area <- st_sfc(st_point(c(25.2967561, 13.5051014)), crs=4326) |>
    st_buffer(100)
```

Let's use our crop footprint..

```{r}
site_2_20220526_cropped <- site_2_20220526 |>
    crop(crop_area |> st_transform(32635))
```

See how it looks.

```{r}
ggplot() + 
    geom_spatraster_rgb(data=site_2_20220526_cropped)
```

Great. Let's write it.

```{r}
writeRaster(site_2_20220526_cropped, "/Users/a0161776/Documents/nicar/site_2_20220526_cropped.tif")
```

## 20 April 2023

Load another image, do pre-processing steps.

```{r}
rast("/Users/a0161776/Documents/nicar/site_2_20230420.tif") |>
    shift(2, -7) |>
    writeRaster("/Users/a0161776/Documents/nicar/site_2_20230420_shifted.tif", overwrite=TRUE)
```

## 28 October 2024

Load another image, do pre-processing steps.

```{r}
rast("/Users/a0161776/Documents/nicar/site_2_20241028.tif") |>
    shift(2, -9) |>
    writeRaster("/Users/a0161776/Documents/nicar/site_2_20241028_shifted.tif", overwrite=TRUE)
```

## Put it all together

Load points.

```{r}
all_points <- st_read("/Users/a0161776/Documents/nicar/all_points.gpkg") |>
    rowid_to_column()
```

Now let's load our polys.

```{r}
site_2_20220526_polys <- st_read("/Users/a0161776/Documents/nicar/site_2_20220526_polys.geojson") |>
    mutate(dt = ymd("20220526"))

site_2_20230420_polys <- st_read("/Users/a0161776/Documents/nicar/site_2_20230420_polys.geojson") |>
    mutate(dt = ymd("20230420"))

site_2_20241028_polys <- st_read("/Users/a0161776/Documents/nicar/site_2_20241028_polys.geojson") |>
    mutate(dt = ymd("20241028"))
```

Now we can combine them into a single SF.

```{r}
all_polys <- site_2_20220526_polys |>
    bind_rows(site_2_20230420_polys) |>
    bind_rows(site_2_20241028_polys)

all_polys
```

Now we can do a spatial join...

```{r}
all_points |>
    st_join(all_polys)
```

Split-apply-combine and all...

```{r}
points_min_dates <- all_points |>
    st_join(all_polys) |>
    st_drop_geometry() |>
    group_by(rowid) |>
    summarise(min_dt = min(dt))
```

Join to give us a min date field...

```{r}
points_min_dates_final <- all_points |>
    left_join(points_min_dates)
```

See how it looks.

```{r}
ggplot() + 
    geom_sf(data=points_min_dates_final, aes(color=min_dt |> as.factor()))
```

Let's do math.

```{r}
points_min_dates_final |>
    st_drop_geometry() |>
    group_by(min_dt) |>
    summarise(tot = n()) |>
    mutate(previous_dt = lag(min_dt)) |>
    mutate(days_between = difftime(min_dt, previous_dt, units = c("days")) |> as.integer()) |>
    mutate(pace = tot / days_between)
```
